\vspace{-2cm} {\bf Online Appendix} \\ Government websites as data: \\ A methodological pipeline with application to the websites of municipalities in the United States

### 2020-01-17 paper/archived/manuscript_ssqNote.tex
revising for APR

%============Article Title, Authors==================
\title{\vspace{-2cm} Government websites as data: \\ A methodological pipeline with application to the websites of municipalities in the United States}

\author{ Markus Neumann\footnote{Department of Political Science, The Pennsylvania State University, University Park, PA 16802, USA. Email: mvn5218@psu.edu. Corresponding author.} \and Fridolin Linder\footnote{Department of Political Science, Social Media and Political Participation Lab, New York University, New York, NY 10012, USA. Email: fridolin.linder@nyu.edu} \and Bruce Desmarais\footnote{Department of Political Science, The Pennsylvania State University, University Park, PA 16802, USA. Email: bdesmarais@psu.edu}} \date{\today}

@@

A local government's website is an important source of information about policies and procedures for residents and community stakeholders. Existing research in political science and related fields has relied on manual methods of website content collection, limiting the scale and scope of website content analysis. Our objectives and corresponding contributions are two-fold. First, we develop a methodological pipeline to gather, process, and analyze website content, along with an R package implementation of the pipeline. Second, we illustrate the pipeline through the collection and analysis of the websites of over two hundred municipal governments in the United States. We build upon recent research that analyzes how variation in the partisan control of government relates to content posted on the government's website.
%\noindent We explore the effect of transitions of power in municipal governments on the content of their websites. We hypothesize that when party control changes, city administrators modify the contents of their websites in order to fit the agenda of the new incumbent. To test this theory, we study cities in Indiana and Louisiana, two states in which all municipal elections are partisan and the parties of the candidates appear on the ballots. Snapshots of websites before and after transitions of power are acquired through the Wayback Machine. We apply statistical topic models based in latent dirichlet allocation, focusing on changes to the websites. We present results on both which topics see the greatest degree of change associated with transitions in city administrations, and how the topics modified differ with regard to political parties.

@@

\section{Introduction}
Government websites convey voluminous information about all aspects of government policymaking, policy implementation, and public deliberation. The vital role of official websites in connecting the government and the governed has motivated a wave of research on the contents of government websites, focusing in particular on textual contents \citep[e.g.,][]{grimmelikhuijsen2010transparency,wang2005evaluating,osman2014cobra,Eschenfelder2002}. The conventional approach to data collection in projects focused on government websites involves manual content extraction from each website in the dataset. Though accurate, the manual approach to data collection is costly for large-scale analysis. We present a methodological pipeline that can be used to automatically scrape government websites in order to build datasets that can be used for text analysis---describing challenges in data collection and processing, as well as the solutions we adopt. We provide an illustrative application in which we explore the ways in which the textual contents on city government websites in six American states correlate with the partisanship of the city mayor.

@@

\section{Politics and Government Website Content}

@@

\begin{figure}
\centering
\includegraphics[scale=0.45]{figures/gary_hp}
\caption{Screenshot from the homepage at \url{https://garyin.us/}, accessed on 05/22/2019. Image depicts Democratic mayor of Gary, IN, Karen Freeman-Wilson.}
\label{fig:garymayor}
\end{figure}

@@

The existing research that uses scraped websites provides an indication of the theoretical value of empirical analysis of web contents. Research on `e-governance' evaluates government websites in terms of accessibility, ease-of-use, and function \citep[e.g., ][]{Urban2002,mcnutt2010virtual,Armstrong2011,Feeney2017}. As an example, \cite{grimmelikhuijsen2012developing} study local government websites of Dutch municipalities to measure government transparency regarding air quality in the municipalities. The websites of politicians and their parties have also been the object of research \citep{Druckman2009,Druckman2010,Cryer2017,Esterling2011,Esterling2011a,Norris2003,Therriault2010}. For example, \cite{Druckman2010} analyze the issues engaged on websites for candidates in U.S. Congressional elections, and find that candidates strategically engage just a few issues based on the priorities in their districts and the characteristics of their opponents.

@@

\section{Data: US Municipal Government Website Text}

@@

\section{The Web to Text Pipeline}
In this section, we describe our methodological pipeline, with which we take an archive of website files, and output a corpus of formatted plain text documents. We address three methodological challenges.  First, though they contain significant amounts of text, websites are not comprised of clean plain text files. Rather, the files available at websites are of multiple types, including HTML, PDF, word processor, plain text, and image files. The first step is aimed at extracting clean plain text from this heterogeneous file base. The second step in our pipeline is to process the text to remove language that is effective at differentiating one website from another but is uninformative regarding policy or political differences between governments. Finally, these tools need to work consistently across all of the websites in our corpus, in spite of the fact that relevant information is stored and structured in different ways. We make a software recommendation for each of these steps and gather most of them our R package, \texttt{gov2text}. All of the recommended software is either well-established in the natural language processing community, or part of the Unix ecosystem. As such, all of it is free, open source, well-developed and will continue to be supported by a dedicated community. Some of the steps we take in this processing pipeline are universally applicable in the analysis of textual data, and some of them are most appropriate for the particular type of text analysis that we apply to this data---statistical topic modeling. We will clarify this distinction as we describe steps in our pipeline.

@@

A subfield of the information retrieval literature, dealing with boilerplate extraction, can offer a solution to this problem. The goal of this branch of research is to develop algorithms with the ability to estimate whether a given portion of an HTML file is substantive. To this end, structural features, such as HTML tags (which are not sufficiently informative on their own), text statistics such as word and sentence length, as well as other heuristics are used. We rely on the \texttt{boilerpipe} classifier described in \cite{Kohlschutter2010}, which is implemented through the R package \texttt{boilerpipeR}. The \texttt{boilerpipe} algorithm has been widely used in the computer science and natural language processing literatures, but to our knowledge has not been previously used in political science.   The complete text extracted from the Gary, IN homepage using boilerpipe is depicted in the screenshot in the bottom row of Figure \ref{fig:garysbs}. We see that only the Mayor's message is extracted, leaving the rest of the text as boilerplate.\footnote{In the online appendix we present a replication of the topic modeling presented in the main text below in which we use a minimal HTML parser rather than \texttt{boilerpipe} to process the data. We show that without \texttt{boilerpipe}, some of the most partisan 'topics' are simply website boilerplate text. }

@@

\begin{table}[ht]
    \centering
    \begin{tabular}{lrr}
        \hline
        Process & Software dependency & in \texttt{gov2text} \\
        \hline
        1. Assemble url list. & \texttt{Selenium} & no \\
        2. Collect website files. & \texttt{wget} & no \\
        3. Correct file extensions. & \texttt{wand} \citep{wand} & yes \\
        4. Discard website boilerplate. & \texttt{boilerpipeR} \citep{boilerpipeR} & yes \\
        5. Convert non-HTML files to text. & \texttt{readtext} \citep{readtext} & yes \\
        6. Lemmatize text. & \texttt{spacyr} \citep{spacyr}& yes \\
        7. Remove names. & \texttt{spacyr} & yes \\
        8. Retain nouns, verbs, adjectives. & \texttt{spacyr} & yes \\
        9. Stopword/number removal. & \texttt{quanteda} \citep{quanteda} & yes \\ 
        10. Retain only English words. & \texttt{Hunspell} \citep{hunspell} & yes \\
        11. Removal of duplicate documents. & \texttt{gov2text} & yes \\
        \hline
    \end{tabular}
    \caption{\label{tab:list} Data collection and processing pipeline. Steps to collect and prepare text for topic modeling.} 
\end{table}

@@

To study content differences between government websites based on mayoral partisanship, we draw upon a recently-developed class model for text, the structural topic model (STM), developed by \citet{Roberts2014}. Building on the conception of ``topics'' in Latent Dirichlet Allocation, in the STM a topic is a multinomial distribution defined on the word types in the corpus dictionary. The log-odds of the topic probabilities in each document-specific multinomial distribution over topics are drawn from a multivariate normal distribution in which the topic-specific means are determined by a linear regression function that associates document-attributed covariates with topics. For example, in the context of municipal website content, the structural topic model can be used to estimate a regression coefficient that defines the linear relationship between the log-odds of the municipality's population and the log-odds of each topic. For our primary empirical investigation, the STM provides a tool to estimate the relationship between the party of the city's mayor and the prevalence of each topic. We also include the municipality's population and median income as covariates. Further details on and results from our STM specification can be found in the online appendix.

@@

\section*{Funding}
This work was supported by the National Science Foundation [1320219, 1637089, 1641047].

>>>>

%============Article Title, Authors==================
\title{\vspace{-2cm} Government Websites As Data: \\ A methodological pipeline with application to the websites of municipalities in the United States}

%\author{ Markus Neumann\footnote{Department of Political Science, The Pennsylvania State University, University Park, PA 16802, USA. Email: mvn5218@psu.edu. Corresponding author.} \and Fridolin Linder\footnote{Department of Political Science, Social Media and Political Participation Lab, New York University, New York, NY 10012, USA. Email: fridolin.linder@nyu.edu} \and Bruce Desmarais\footnote{Department of Political Science, The Pennsylvania State University, University Park, PA 16802, USA. Email: bdesmarais@psu.edu. This work was supported by the National Science Foundation [1320219, 1637089, 1641047].}} \date{\today}

@@

\noindent {\bf Objective:} Existing social science research that uses government website content has relied on manual methods of content collection, limiting the scale and scope of projects. Our objectives and corresponding contributions in this research note are two-fold. First, we develop a methodological pipeline to gather, process, and analyze website content, along with an R package implementation of the pipeline. Second, we illustrate the pipeline through the collection and analysis of the websites of over two hundred municipal governments in the United States. 
\vspace{.1cm} \\
\noindent {\bf Methods:} We propose and utilize a pipeline consisting of tools for web-scraping, text parsing and cleaning, and text analysis.  \vspace{.1cm} \\
\noindent {\bf Results:} Our pipeline is effective in extracting large-scale meaningful substantive text from government websites. We identify a strong relationship between the topical content on municipal websites and the partisanship of the mayor.  \vspace{.1cm} \\
\noindent {\bf Conclusion:}  The pipeline we develop, associated software, and dataset of municipal websites, represent valuable research tools for social scientists.  \vspace{-.3cm}

%\noindent We explore the effect of transitions of power in municipal governments on the content of their websites. We hypothesize that when party control changes, city administrators modify the contents of their websites in order to fit the agenda of the new incumbent. To test this theory, we study cities in Indiana and Louisiana, two states in which all municipal elections are partisan and the parties of the candidates appear on the ballots. Snapshots of websites before and after transitions of power are acquired through the Wayback Machine. We apply statistical topic models based in latent dirichlet allocation, focusing on changes to the websites. We present results on both which topics see the greatest degree of change associated with transitions in city administrations, and how the topics modified differ with regard to political parties.

@@

\section{Introduction} \vspace{-.2cm}
Government websites convey voluminous information about all aspects of government policymaking, policy implementation, and public deliberation. The vital role of official websites in connecting the government and the governed has motivated a wave of research on the contents of government websites, focusing in particular on textual contents \citep[e.g.,][]{grimmelikhuijsen2010transparency,wang2005evaluating,osman2014cobra,Eschenfelder2002}. The conventional approach to data collection in projects focused on government websites involves manual content extraction from each website in the dataset. Though accurate, the manual approach to data collection is costly for large-scale analysis. We present a methodological pipeline that can be used to automatically scrape government websites in order to build datasets that can be used for text analysis---describing challenges in data collection and processing, as well as the solutions we adopt. We provide an illustrative application in which we explore the ways in which the textual contents on city government websites in six American states correlate with the partisanship of the city mayor.

@@

\vspace{-.2cm}
\section{Politics and Government Website Content} \vspace{-.2cm}

@@

\begin{figure}
\centering
\includegraphics[scale=0.45]{figures/gary_hp}
\caption{Screenshot from the homepage at \url{https://garyin.us/}, accessed on 05/22/2019. Image depicts Democratic mayor of Gary, IN, Karen Freeman-Wilson.} \vspace{-.3cm}
\label{fig:garymayor}
\end{figure}

@@

The existing research that uses scraped websites provides an indication of the theoretical value of empirical analysis of web contents. Research on `e-governance' evaluates government websites in terms of accessibility, ease-of-use, and function \citep[e.g., ][]{Urban2002,mcneal2003innovating,tolbert2008institutions,mcnutt2010virtual,Armstrong2011,Feeney2017}. As an example, \cite{grimmelikhuijsen2012developing} study local government websites of Dutch municipalities to measure government transparency regarding air quality in the municipalities. The websites of politicians and their parties have also been the object of research \citep{Druckman2009,Druckman2010,Cryer2017,Esterling2011,Esterling2011a,Norris2003,Therriault2010}. For example, \cite{Druckman2010} analyze the issues engaged on websites for candidates in U.S. Congressional elections, and find that candidates strategically engage just a few issues based on the priorities in their districts and the characteristics of their opponents.

@@

\vspace{-.2cm}

\section{Data: US Municipal Government Website Text} \vspace{-.2cm}

@@

\vspace{-.2cm}
\section{The Web to Text Pipeline} \vspace{-.2cm}
In this section, we describe our methodological pipeline, with which we take an archive of website files, and output a corpus of formatted plain text documents. We address three methodological challenges.  First, though they contain significant amounts of text, websites are not comprised of clean plain text files. Rather, the files available at websites are of multiple types, including HTML, PDF, word processor, plain text, and image files. The first step is aimed at extracting clean plain text from this heterogeneous file base. The second step in our pipeline is to process the text to remove language that is effective at differentiating one website from another but is uninformative regarding policy or political differences between governments. Finally, these tools need to work consistently across all of the websites in our corpus, in spite of the fact that relevant information is stored and structured in different ways. We make a software recommendation for each of these steps and gather most of them our R package, \texttt{gov2text}. All of the recommended software is either well-established in the natural language processing community, or part of the Unix ecosystem. As such, all of it is free, open source, well-developed and will continue to be supported by a dedicated community. Some of the steps we take in this processing pipeline are universally applicable in the analysis of textual data, and some of them are most appropriate for the particular type of text analysis that we apply to this data---statistical topic modeling. We will clarify this distinction as we describe steps in our pipeline.

@@

A subfield of the information retrieval literature, dealing with boilerplate extraction, can offer a solution to this problem. The goal of this branch of research is to develop algorithms with the ability to estimate whether a given portion of an HTML file is substantive. To this end, structural features, such as HTML tags (which are not sufficiently informative on their own), text statistics such as word and sentence length, as well as other heuristics are used. We rely on the \texttt{boilerpipe} classifier described in \cite{Kohlschutter2010}, which is implemented through the R package \texttt{boilerpipeR}. The \texttt{boilerpipe} algorithm has been widely used in the computer science and natural language processing literatures, but to our knowledge has not been previously used in the social sciences.   The complete text extracted from the Gary, IN homepage using boilerpipe is depicted in the screenshot in the bottom row of Figure \ref{fig:garysbs}. We see that only the Mayor's message is extracted, leaving the rest of the text as boilerplate.\footnote{In the online appendix we present a replication of the topic modeling presented in the main text below in which we use a minimal HTML parser rather than \texttt{boilerpipe} to process the data. We show that without \texttt{boilerpipe}, some of the most partisan 'topics' are simply website boilerplate text. }

@@

\begin{table}[ht]
    \centering
    \begin{tabular}{lrr}
        \hline
        Process & Software dependency & in \texttt{gov2text} \\
        \hline
        1. Assemble url list. & \texttt{Selenium} & no \\
        2. Collect website files. & \texttt{wget} & no \\
        3. Correct file extensions. & \texttt{wand} \citep{wand} & yes \\
        4. Discard website boilerplate. & \texttt{boilerpipeR} \citep{boilerpipeR} & yes \\
        5. Convert non-HTML files to text. & \texttt{readtext} \citep{readtext} & yes \\
        6. Lemmatize text. & \texttt{spacyr} \citep{spacyr}& yes \\
        7. Remove names. & \texttt{spacyr} & yes \\
        8. Retain nouns, verbs, adjectives. & \texttt{spacyr} & yes \\
        9. Stopword/number removal. & \texttt{quanteda} \citep{quanteda} & yes \\ 
        10. Retain only English words. & \texttt{Hunspell} \citep{hunspell} & yes \\
        11. Removal of duplicate documents. & \texttt{gov2text} & yes \\
        \hline
    \end{tabular}
    \caption{\label{tab:list} Data collection and processing pipeline. Steps to collect and prepare text for topic modeling.}  \vspace{-.9cm}
\end{table}

@@

To study content differences between government websites based on mayoral partisanship, we draw upon a recently-developed class model for text, the structural topic model (STM), developed by \citet{Roberts2014}. Building on the conception of ``topics'' in conventional topic models \citep{valdez2018topic}, in the STM a topic is a multinomial distribution defined on the word types in the corpus dictionary. The log-odds of the topic probabilities in each document-specific multinomial distribution over topics are drawn from a multivariate normal distribution in which the topic-specific means are determined by a linear regression function that associates document-attributed covariates with topics. For example, in the context of municipal website content, the structural topic model can be used to estimate a regression coefficient that defines the linear relationship between the log-odds of the municipality's population and the log-odds of each topic. For our primary empirical investigation, the STM provides a tool to estimate the relationship between the party of the city's mayor and the prevalence of each topic. We also include the municipality's population and median income as covariates. Further details on and results from our STM specification can be found in the online appendix.

@@

### 2020-01-17 paper/archived/manuscript_ssqNote_OA copy.tex
revising for APR

\author{ Markus Neumann\footnote{Department of Political Science, The Pennsylvania State University, University Park, PA 16802, USA. Email: mvn5218@psu.edu. Corresponding author.} \and Fridolin Linder\footnote{Department of Political Science, Social Media and Political Participation Lab, New York University, New York, NY 10012, USA. Email: fridolin.linder@nyu.edu} \and Bruce Desmarais\footnote{Department of Political Science, The Pennsylvania State University, University Park, PA 16802, USA. Email: bdesmarais@psu.edu}} \date{\today}

@@

A local government's website is an important source of information about policies and procedures for residents, community stakeholders and scholars. Existing research in public administration, public policy, and political science has relied on manual methods of website content collection and processing, limiting the scale and scope of website content analysis. We develop a methodological pipeline that researchers can follow in order to gather, process, and analyze website content. Our approach, which represents a considerable improvement in scalability, involves downloading the entire contents of a website, extracting the text and discarding redundant information. We provide an R package that can be used to apply our proposed pipeline. We illustrate our methodological pipeline through the collection and analysis of a new and innovative dataset---the websites of over two hundred municipal governments in the United States. We build upon recent research that analyzes how variation in the partisan control of government relates to content made available on the government's website. Using a structural topic model, we find that cities with Democratic mayors provide more information on policy deliberation and crime control, whereas Republicans prioritize basic utilities and services such as water, electricity and fire safety.
%\noindent We explore the effect of transitions of power in municipal governments on the content of their websites. We hypothesize that when party control changes, city administrators modify the contents of their websites in order to fit the agenda of the new incumbent. To test this theory, we study cities in Indiana and Louisiana, two states in which all municipal elections are partisan and the parties of the candidates appear on the ballots. Snapshots of websites before and after transitions of power are acquired through the Wayback Machine. We apply statistical topic models based in latent dirichlet allocation, focusing on changes to the websites. We present results on both which topics see the greatest degree of change associated with transitions in city administrations, and how the topics modified differ with regard to political parties.

>>>>

%\author{ Markus Neumann\footnote{Department of Political Science, The Pennsylvania State University, University Park, PA 16802, USA. Email: mvn5218@psu.edu. Corresponding author.} \and Fridolin Linder\footnote{Department of Political Science, Social Media and Political Participation Lab, New York University, New York, NY 10012, USA. Email: fridolin.linder@nyu.edu} \and Bruce Desmarais\footnote{Department of Political Science, The Pennsylvania State University, University Park, PA 16802, USA. Email: bdesmarais@psu.edu}} \date{\today}

@@

\noindent {\bf Objective:} Existing social science research that uses government website content has relied on manual methods of content collection, limiting the scale and scope of projects. Our objectives and corresponding contributions in this research note are two-fold. First, we develop a methodological pipeline to gather, process, and analyze website content, along with an R package implementation of the pipeline. Second, we illustrate the pipeline through the collection and analysis of the websites of over two hundred municipal governments in the United States. 
\\~\\
\noindent {\bf Methods:} We propose and utilize a pipeline consisting of tools for web-scraping, text parsing and cleaning, and text analysis.  \\~\\
\noindent {\bf Results:} Our pipeline is effective in extracting large-scale meaningful substantive text from government websites. We identify a strong relationship between the topical content on municipal websites and the partisanship of the mayor.  \\~\\
\noindent {\bf Conclusion:}  The pipeline we develop, associated software, and dataset of municipal websites, represent valuable research tools for social scientists.

%\noindent We explore the effect of transitions of power in municipal governments on the content of their websites. We hypothesize that when party control changes, city administrators modify the contents of their websites in order to fit the agenda of the new incumbent. To test this theory, we study cities in Indiana and Louisiana, two states in which all municipal elections are partisan and the parties of the candidates appear on the ballots. Snapshots of websites before and after transitions of power are acquired through the Wayback Machine. We apply statistical topic models based in latent dirichlet allocation, focusing on changes to the websites. We present results on both which topics see the greatest degree of change associated with transitions in city administrations, and how the topics modified differ with regard to political parties.
